Identifying 15-minute neighbourhoods with complete caring potential

### 2024-09-30 paper/Manuscript.qmd
draft of Abstract through Methods

The "15-Minute City" has been recently adopted by leaders as a way to promote human-scaled cities (e.g., Paris mayor Anne Hidalgo proposed a 15-minute city plan in her 2020 re-election campaign, 15-minute walkable neighbourhood guidance in Shanghai, China [@weng15minuteWalkableNeighborhoods2019]), As defined in @moreno_introducing_2021, the 15-Minute City is a urban planning model based in chrono-urbanism, a theory that emphasizes the positive impact on quality of urban life when urban space becomes multi-rythmic, in opposition to industrial Fordism, single-use zoning, and the segregation of urban time and space for individual uses and mobility [@mulicekUrbanRhythmsChronotopic2015; @morenoVilleQuartHeure2016]. The 15-Minute City is one that accommodates the ability to reach essential destinations within a walkable, bikeable or transit supportive radius. It would allow individuals to reclaim time spent on car mobility and give way to sustainable modes; prompting urban spaces that are responsive to human needs and environmental sensibilities [@Allam2022].

Though the term 15-Minute City is relatively new, neighbourhood planning practice aiming to foster community and local travel to amenities is not. As recent examples, planned neighbourhoods dominated post-WWII built urban form: the Neighbourhood Unit Concept in the western world [@brodyNeighbourhoodUnitConcept2013] and a parallel concept of the “mikrorayons” neighbourhoods in the Soviet Union  [@kissfazekasCircleParadigms15minute2022]. These planning forms have also been extensively critiqued: as summarized in @talenSocialSciencePlanned2017, urban planners struggled to determine suitably universal design aspects such as neighborhood size, the type and quality of amenities, and local and regional connectivity. Social scientists also emphasized that top-down neighbourhood planning approaches overestimated the built environment's influence on social life [@talenSocialSciencePlanned2017]. Neighbourhood planning approaches such as the 15-Minute City concept are being re-embraced, but from bottom-up perspective. However how to equitably prescribe urban form that 'enables' contact with social opportunities (instead of 'engendering it' from the top-down) is still under debate. For instance, questions such as _what_ are destinations that matter, by what mode and travel time threshold, and _for whom_ remain under discussion among proximity-based planners [@silvaProximitycentredAccessibilityConceptual2023; @silvaRegionalAccessibilityUndermining2022].

To add clarity to these questions surrounding the 15-Minute City concept, in this work we aim to build a theoretical bridge to "Mobility of Care" as introduced in @sanchezdemadariagaMobilityCareIntroducing2013. Mobility of Care is especially relevant to the 15-Minute City, as it is a purpose of travel that is necessary for all and often occurs at the local level. Namely, Mobility of Care refers to all travel required to sustain the needs of a household, such as grocery shopping, escorting children, travelling to health appointments, and running errands. While decades of research have examined types of household-serving trips (e.g., shopping trips, escorting to school), @sanchezdemadariagaMobilityCareIntroducing2013 was the first to consider them all as one category. In doing so, Mobility of Care is shown to be a significant proportion of daily travel, characterizing the trip purpose of approximately 30% of adults’ trips [@sanchezdemadariagaMobilityCareIntroducing2013; @sanchezdemadariagaMeasuringMobilitiesCare2019; @ravensbergen2023exploratory]. Notably, care trips take on tremendous gendered and socio-economic importance, as women, and especially those from lower income households, complete the majority of care trips [@ravensbergen2023exploratory]. Despite its importance for all and in advancing gender equality, the examination of mobility of care and access to care destinations have been under-examined relative to employment destinations. This is especially pertinent in accessibility research that has largely focused on employment () and retail ().

We aim to theoretically bridge Mobility of Care with the 15-Minute City concept to clarify questions surrounding _what destinations within a 15 minute walkable radius matter_. We hypothesis that _care_ destinations matter, they matter for all people but especially for women and lower-income women preforming care trips. In this vein, this work pairs this hypothesis with an exploratory clustering methodology to identify which neighbourhoods have the potential to support 15-minute access to care destinations in an empirical case study of Hamilton, Ontario, typical mid-sized Canadian city. More concretely, the research objectives are two-fold: after a review of the 15-Minute City concept and associated measurement tools, this work first introduces the Mobility of Care conceptualisation to re-imagine local amenities importance within the 15-Minute City concept, and second to apply a data-driven classification and clustering machine learning methodology to support the bottom-up examination of _15-minute caring neighbourhoods_  through an empirical examination of Hamilton.

@@

In the last decades, the need for more sustainable, healthier, and liveable cities have received more attention. Planners and decision-makers have proposed alternatives to mobility-based neighbourhood planning to one centered on proximity to urban functions [@pozoukidou15MinuteCityDecomposing2021]. In this context, the 15-Minute City is now in the public spotlight [@logan_xminute_2022; @moreno_introducing_2021]. In @moreno_introducing_2021 where the 15-Minute City concept was introduced, four universal dimensions are outlined: density (in terms of people per km$^2$), diversity (including mixed-land use and diversity of people), the temporal and spatial proximity to essential services, <!-- (residential areas, work, commercial areas, education centres, health facilities and other basic institutions)--> and digitalisation. The framework presented has been critised within academic and planning circles [@guzmanProximityEnoughCritical2024]. Namely, critiques have centered on the concept's inadequecy in addressing pre-existing structural forces and individual charctertes that drive inequalities in who does and who could benefit more from such a paradigm [@dimarino15minuteCityConcept2023; @willberg15minuteCityAll2023]. A 15-Minute City for all people is a utopic ideal, however how will existing inequalities in how people move through and benefit from the city be addressed. _Who is in the most need of the 15-Minute City before we all get there?_. Without directed answers to these context-sensitive questions makes this increasingly popular concept susceptible to becoming a place-branding slogan [@pozoukidou15MinuteCityDecomposing2021].

Reflecting this flexible framework; cities that have adopted 15-Minute City concepts have done so using a diverse range of definitions and tools along with similar indistinct universal approaches. An early adopter was Portland, US in the Portland Plan [@city_of_portland_20minute_2010] of April 2012. It aims to foster an inclusive city development based on prosperity, education, health, and equity for a 30 year horizon. It promotes neighbourhood self-suffiency and connectivity to city centers and centres of employment. The progress report describes a high-level focus on equitable service delivery to all residents with equity concerning topics related to racial equity and people with disabilities [@portland_government_portland_2017]; similarly adopting an "all populations" approach. Later, other cities also adopted proximity-based goals into their plans, taking similar neutral entity approaches. For instance, in the plan that later catapulted the 15-Minute City into public discourse: Paris mayor Anne Hidalgo proposed a 15-minute city plan in her 2020 re-election campaign that foregrounded six social functions of the city that should be easily accessible from any location, namely: housing, work, health care, groceries, education and leisure [@ville_de_paris_paris_2022]. These ideas inspired language in the agendas of other cities in the Western world such as Ottawa, in the Canadian context, who also adopted a 15-Minutes approach in their recent Official Plan [@ville_dottawa_quartier_2021]. From the review of worldwide practice literature in @teixeiraClassifying15minuteCities2024, overall the 15-Minute City concept is in early phases of implementation around the world and the diverse range of definitions, strategies, and instruments present a significant knowledge and implementation gap: the 15-Minute City seems nice but _how do we get there and whom will benefit_?

The past can shed light on cautionary tales. While the 15-Minute City concept is new, neighbourhood planning to improve society outcomes is not: in fact the literature has drawn parallels from the 15-Minute City to Clarence A. Perry's Neighbourhood Unit Concept (NUC) from the 1930s [@kissfazekasCircleParadigms15minute2022]. The NUC is a socio-spatial normative scheme widely adopted by government agencies (and the real-estate community) in the Western world after the Second World War [@talenSocialSciencePlanned2017; @solow1969concept]. Pairing well with the objectives of planning agencies at the time, Perry's NUC would allow for efficient mass-building of cellular units that prioritized the perceived functional needs of women and children: each unit providing proximate access to an elementary school and sufficent supporting community facilities, shopping, parks and housing [@talenSocialSciencePlanned2017; @brodyNeighbourhoodUnitConcept2013]. The NUC primarily prioritized local service provision, though Perry had confidence in good design's contribution to 'neighbourhood spirit' [@hall2014cities]. From the NUC's conception until the end of the 1960s, planner's attempts to prescribe social meaning to the neighbourhood's physical form had been criticized to near extinction [@talenSocialSciencePlanned2017]. By social scientists, the overestimation on the impact of built environment on social life was critiqued. By other planners, the inability to agree on the specificity of the neighbourhood (i.e., population sized, and the type, quality, and quantity of amenity) as well as how neighbourhood units are connected to other neighbourhoods and the rest of the region.  In response to these criticisms, neighbourhood planning proponents have redefined deterministic terminology: prescribed physical forms as 'enabling' rather than engendering, and the provision of social contact 'opportunities' instead of determined social behaviour. The re-defined bottom-up approach to community and neighbourhood planning has been adopted by New Urbanist from which the 15-Minute City stems [@kissfazekasCircleParadigms15minute2022]. But with _what bottom-up processes?_; small scale experiments are being tested () with preliminary small-scacle results. These processes are only in early stages so critiques of social scientists and planners from decades past still persist.

_How can physical form be planned to enable an improved quality of life, and for whom?_  has also occupied urban planning research in recent years, exploring what are the impacts of built environment on residence's outcomes. For instance, the examination of low-income households access to transportation and their likelihood of gaining employment [@blumenbergDriveWorkRelationship2017; @bastiaanssenDoesBetterJob2022] or the relationship between children's access to public transit and participation in after-school activities [@palmRolePublicTransit2020]. Relatedly, a new wave of researchers and practitioners focused on local and context-specific relationships in the proximity to destinations have also emerged [@silvaProximitycentredAccessibilityConceptual2023; @silvaRegionalAccessibilityUndermining2022]. As reviewed  in the city plans that have adopted 15-Minute City approaches, the NUC's critisms, and recent proximity-based examinations, the question of _how to enable improved quality of life through urban built environment_ has shown to be highly context-senstive with the need for further investigation.

@@

In examining how to enable improved quality of life through urban built environment, an increasingly popular tool have been accessibility measures. They are a way to quantify the ease of reaching destinations from a given point in space and have been used to examine urban areas through just and sustainable city planning agendas [@vale2023accessibility; @handyAccessibilityIdeaWhose2020]. As the 15-Minute City is an amenity-provision neighbourhood planning concept which aims to enable urban environment that improves life quality, its analysis lends well to the use of accessibility tools [@guzmanProximityEnoughCritical2024]. Foreseeably, recent works have applied accessibility measures to investigate the 15-Minute City across various geographic scopes. For instance, in Naples, Italy [@gaglioneUrbanAccessibility15minute2022], Barcelona, Spain [@graellsCityCitiesMeasuring2021], Vancouver, Canada [@hosford15minuteCityReach2022], and in urban areas across Europe [@vale2023accessibility]. The "cumulative opportunity" measure has been applied to many 15-Minute City accessibility examinations. This form of accessibility quantifies how many destinations can be reached from a point in space within a given travel time threshold, pairing well with examinations that have a normative travel time threshold, like the 15-Minute City. Furthermore, the cumulative opportunity measure is widely appreciated for its intuitive computation and is popular among practitioners [@handyAccessibilityIdeaWhose2020; @handyMeasuringAccessibilityExploration1997; @chengInvestigatingWalkingAccessibility2019]. However, reflecting the diversity of the accessibility literature, accessibility measures other than the cumulative opportunity have also been applied [@guzmanProximityEnoughCritical2024].

Another method which compliments the assessment of the 15-Minute City as it relates to the diversity of opportunity types, is entropy. The entropy measure has been widely used to characterise the diversity of land-use mix in early work by @frankLinkingObjectivelyMeasured2005 based on the species diversity Shannon-Wiener index that expresses relative evenness throughout a sequence (this measure among other species diversity indices of the time are reviewed in @whittaker1972evolution). It has also been used in social sciences to comprehend the occurrence of social events in a structured manner specially with regard to life trajectories [@ritschardSequenceAnalysisRelated2018].In the realm of urban studies and planning literature the malleable concept of entropy has been used to characterise the diversity in land-use mix [@ewingTravelBuiltEnvironment2010a], especially in the context of active modes [@luUrbanDensityDiversity2017; @mavoaIdentifyingAppropriateLanduse2018] and suburban sprawl [@randallGISBasedLand2015]. As well, it has been used to understand mobility behaviour [@mcbride2020exploration; @montero2023applying; @montero2023role]. Work that has used diversity indices alongside accessibility analysis are scarce, though examples of its use as parameters within accessibility scores in the case of employment opportunities [@chengMeasuringUrbanJob2013; @daiIncorporatingJobDiversity2018] and more recently to describe the diversity in transit facilities [@yinIncorporatingFacilityDiversity2024] are present. To the authors knowledge, the use of an entropy measure to reflect the diversity in care destination amenity types and in context within a 15-minute neighbourhood, has yet to be done. However, entropy measures thematically compliments the 15-Minute City.

`r word_comment("I need to fill this out. A small classification methods paragraph", ".")`
In addition, a useful tool has been classification algorithms in the identification of spatial typologies and in profiling who may reside there. Those too have K means... Hierarchical... and SOM [@delmelleDifferentiatingPathwaysNeighborhood2017]... regionlization techniques such as spatially constrained multivariate clustering [@burkeGeospatialAnalysisFramework2022] Use of tree diagrams! ....SOM is a type of artificial neural network (ANN) trained using unsupervised learning that produces a two-dimensional representation of the input space. It applies competitive learning by way of a neighbourhood function thereby preserving the topological properties of the input space and producing combinations of the input space that we can then intuitively profile. It has been used widely in XX but scarcely in transportation related research [@victorianoTimeSpaceMoney2020]

All-in-all, within the context of examining the 15-Minute City there has been a wide range of tools applied in the literature that appear appriorpiate. However, the question of _15-minute cities for whom_ requires further examination. In this way, this work connects the 15-Minute City to another normative conceptualisation: Mobility of Care.

@@

As the 15-Minute City is in early stages of conceptualisation; an alternative way to conceive of _what_ urban functions matter could be through a feminist lens. Instead of focusing on _all_ destinations, it may be useful to examine destinations involved in caring activities. Caring activities meet the physical, psychological, and emotional needs of others, and are one of the most essential and fundamental activities in society [@ilo_care_2018]. Yet, they are the most unequal, undervalued, and even devalued activities worldwide. Traditionally, caring activities have been borne on women’s shoulders [@haydenGrandDomesticRevolution1982; @hochschild_second_2012]. According to @ilo_care_2018, women and girls perform more than three-quarters of the total amount of unpaid care, a gender gap that varies geographically [@RN5]. This unequal share of caring responsibilities leads to multifaceted gendered differences: in career development, profession selection, contract type, pay gap, and time poverty, as recognized by various international organisations [@EIGE2016; @ilo_care_2018]. In terms of spatial and transportation planning, almost one third of daily trips are for care purposes [@sanchezdemadariagaMobilityCareIntroducing2013; @sanchezdemadariagaMeasuringMobilitiesCare2019; @ravensbergen2023exploratory]. From this research motivation, Sánchez de Madariaga coined the term Mobility of Care in 2013 to refer to all travel required to sustain the needs of a household, such as grocery shopping, escorting children, travelling to health appointments, and running errands [@sanchezdemadariagaMobilityCareIntroducing2013]. While decades of research had examined these unique household-serving trips, her seminal work was the first to consider all these trips as one category and demonstrate how mobility of care is a significant proportion of daily travel.

The adoption of the Mobility of Care concept within the 15-Minute City also explicitly incorporates equity considerations. Perhaps unsurprisingly given the gendered division of care work discussed above, women have consistently been found to complete more mobility of care trips than men. In one study, mobility of care comprised 32% of women’s daily trips compared to 28% of men’s. While this gap was significant, it was found to be far greater in lower income households where women complete 20% more care trips than men [@ravensbergen2023exploratory]. Sánchez de Madariaga not only shows how important these mobility of care trips are, but also highlights the ways in which “mobility of care is systematically under-represented in any analysis of urban transport” (p. 37). Transport surveys and tools do not directly capture mobility of care, which re-enforces the idea that these trips are not a significant part of daily mobility. In this respect, the feminist perspective of the cities of proximity is still underestimated with only few examples on the topic [@gil_sola_choose_2022; @macintyre_her_2022]. Given the impact of caring activities on everyday life, the chrono-urbanism should specifically target the analysis of caring activities.

Mobility of Care also explicitly prioritizes shorter-distance and local trips, those that could be done by sustainable modes if sufficently supported, making the perspective highly compatible with the 15-Minute City paradigm.Compared to the trip to work, mobility of care trips are more frequently completed by foot, and less frequently done by transit or bicycle [@ravensbergen2023exploratory]. Further, evidence suggests that walking is more common for care than for work trips because care trips tend to be short [@ravensbergen2023exploratory]; if care destinations are closer to the home, it is more likely that these destinations are walkable. Ample literature examines the benefits and barriers to walking care destinations, such as schools (e.g., [@RN7;@RN8;@RN9]) and grocery stores (e.g., [@RN10;@RN11]). However, little of this work uses the mobility of care framework. Instead, they tend to consider singular care destinations in research focusing on walkability.Furthermore, mobility of care trip characteristics differ from other trips, notably those of the commute to work. For instance, many of these trips are made through trip-chains, i.e., when a non-work-related stop is made during the commute. Women have been found to do so more than men, which is understood as being due to their need to balance unpaid-household serving care and paid employment (e.g., picking up groceries or children on the way home from work) [@ravensbergen2023exploratory; @scheinerWomenComplexDaily2015].

`r word_comment("I need to clean this up and firmly reiterate the goals", " ")`
Taken together, city planning is moving to a people-centered approach focusing on accessibility to basic needs. However, the 15-Minute City has faced criticism as it fails to provide sufficient guidance on how to reach the utopic goal of 15-Minute for all. We argue, that significance in what destinations matter within the 15-Minute City could be gained by adopting a Mobility of Care prespective. We develop this approach using a mid-size typical Canadian city of Hamilton and a novel care destination dataset as case study. The objectives of the paper are as follows:

>>>>

The "15-Minute City" has been recently adopted by leaders as a way to promote human-scaled cities [@teixeiraClassifying15minuteCities2024; @teixeiraClassifying15minuteCities2024]. As introduced in @moreno_introducing_2021, the 15-Minute City is a urban planning model based in chrono-urbanism, a theory that emphasizes the positive impact on quality of urban life when urban space becomes multi-rythmic, in opposition to the segregation of urban time and space for individual uses and mobility [@mulicekUrbanRhythmsChronotopic2015; @morenoVilleQuartHeure2016]. It is in opposition to the forces that supported the formation of single-use neighbourhoods such as industrial Fordism and single-use zoning [@mulicekUrbanRhythmsChronotopic2015; @morenoVilleQuartHeure2016]. The "15-Minute City" is a utopic ideal, it is a city where one that accommodates the ability for all to reach important destinations within a walkable, bikeable or transit supportive radius. This form of neighbourhood planning would allow individuals to reclaim time spent on car mobility, giving way to sustainable modes and prompting urban spaces that are responsive to human needs and environmental sensibilities [@Allam2022].

Though the term 15-Minute City is relatively new, neighbourhood planning practice aiming to foster community and local travel to amenities is not. As recent examples, planned neighbourhoods dominated post-WWII built urban form: the Neighbourhood Unit Concept in the western world [@brodyNeighbourhoodUnitConcept2013] and a parallel concept of the “mikrorayons” neighbourhoods in the Soviet Union  [@kissfazekasCircleParadigms15minute2022]. These planning forms have also been extensively critiqued: as summarized in @talenSocialSciencePlanned2017, urban planners struggled to determine suitably universal design aspects such as neighborhood size, the type and quality of amenities, and local and regional connectivity. Social scientists also emphasized that top-down neighbourhood planning approaches overestimated the built environment's influence on social life [@talenSocialSciencePlanned2017]. Neighbourhood planning approaches such as the 15-Minute City concept are being re-embraced by some, but from bottom-up perspective. How to equitably prescribe urban form that 'enables' contact with social opportunities (instead of 'engendering it' from the top-down) is still under debate. For instance, questions such as _what_ are destinations that matter, by what mode and travel time threshold, and _for whom_ remain under discussion among proximity-based planners who frequently use accessibility-based tools [@silvaProximitycentredAccessibilityConceptual2023; @silvaRegionalAccessibilityUndermining2022; @guzmanProximityEnoughCritical2024].

To clarify questions surrounding the 15-Minute City concept of what destinations matter and for whom, this work builds a theoretical bridge to "Mobility of Care". The Mobility of Care term was coined in @sanchezdemadariagaMobilityCareIntroducing2013; it refers to all travel required to sustain the needs of a household, such as grocery shopping, escorting children, travelling to health appointments, and running errands.  While decades of research have examined types of household-serving trips (e.g., shopping trips, escorting to school), @sanchezdemadariagaMobilityCareIntroducing2013 was the first to consider them all as one category. In doing so, Mobility of Care is shown to be a significant proportion of daily travel, characterizing the trip purpose of approximately 30% of adults’ trips [@sanchezdemadariagaMobilityCareIntroducing2013; @sanchezdemadariagaMeasuringMobilitiesCare2019; @ravensbergen2023exploratory]. We argue Mobility of Care is especially relevant to the 15-Minute City, as care trips are necessary for _all_ people and they most often occur at the local level. Moreover, care trips take on tremendous gendered and socio-economic importance, as women, and especially those from lower income households, complete the majority of care trips [@ravensbergen2023exploratory]. Despite care travel's importance in advancing gender equality and equity broadly, the examination of mobility of care and access to care destinations have been under-examined relative to employment destinations. This is especially pertinent in accessibility research that has largely focused on employment points of interest e.g., [@farberOntarioLineSocioeconomic2019; @duarteInfluenceJobAccessibility2023; @ryanAccessibilitySpaceTime2023].

The objectives of this research are two-fold. First, we aim to theoretically bridge Mobility of Care with the 15-Minute City concept to to re-imagine what local amenities matter and define an associated accessibility and diversity of accessibility measure. Second, taking a bottom-up data-driven approach, we apply a classification and clustering machine learning methodologies to support the bottom-up identification of _15-minute caring neighbourhoods_  through an empirical examination of the mid-cited city of Hamilton, Canada. These neighbourhoods are then interpreted on the continuum of having (and not having) the region-relative land-use and transportation infrastructure to support _completely caring_ access. Who currently resides in which neighbourhood is examined, and ways to further expand the care accessibility story are discussed.

@@

In the last decades, the need for more sustainable, healthier, and liveable cities have received signficant attention. Planners and decision-makers have proposed alternatives to mobility-based neighbourhood planning principles to ones that center proximity to urban functions [@pozoukidou15MinuteCityDecomposing2021]. In this context, the 15-Minute City is now in the public spotlight [@logan_xminute_2022; @moreno_introducing_2021]. In @moreno_introducing_2021 where the 15-Minute City concept was introduced, four universal dimensions are outlined: density (in terms of people per km$^2$), diversity (including mixed-land use and diversity of people), the temporal and spatial proximity to essential services, <!-- (residential areas, work, commercial areas, education centres, health facilities and other basic institutions)--> and digitalisation. However, the framework presented has been criticized within academic and planning circles [@guzmanProximityEnoughCritical2024]. Critiques highlight the concept falls short in addressing the  pre-existing structural forces and individual characteristics that drive inequalities, influencing who benefits or could benefit the most from such a paradigm [@dimarino15minuteCityConcept2023; @willberg15minuteCityAll2023]. A 15-Minute City for all people is an aspirational ideal, but how will existing mobility and accessibility inequalities be addressed. Without directed and context-specific solutions, this increasingly popular concept risks becoming a place-branding slogan [@pozoukidou15MinuteCityDecomposing2021].

`r word_comment("Lucia, do I correctly reflect the plans you reviewed?", " ")`
Reflecting the flexible and aspirational presentation of the 15-Minute City; cities that have adopted the concept have done so using a diverse range of definitions and tools along with indistinct universal approaches. An early adopter was Portland, US in the Portland Plan [@city_of_portland_20minute_2010] of April 2012. It aims to foster an inclusive city development based on prosperity, education, health, and equity over a 30 year horizon. It promotes neighbourhood self-sufficiency and connectivity to city centers and centres of employment. The progress report describes a high-level focus on equitable service delivery to all residents with equity concerning topics related to racial equity and people with disabilities [@portland_government_portland_2017], similarly taking on an "all populations" approach. Later, other cities also incorporated proximity-based goals into their plans, adopting similar neutral entity approaches. For instance, in the plan that later catapulted the 15-Minute City into public discourse: Paris mayor Anne Hidalgo proposed a 15-minute city plan in her 2020 re-election campaign that foregrounded six social functions that should be easily accessible from any location, namely: housing, work, health care, groceries, education and leisure [@ville_de_paris_paris_2022]. These ideas inspired language in the agendas of other cities in the Western world such as Ottawa, in the Canadian context, who also adopted a 15-Minutes approach in their recent Official Plan [@ville_dottawa_quartier_2021]. From the review of worldwide practice literature in @teixeiraClassifying15minuteCities2024, the 15-Minute City concept is in early phases of implementation around the world and the diverse range of definitions, strategies, and instruments present a significant knowledge and implementation gap: the 15-Minute City is aspirational but _how do we get there and whom will benefit_?

The past can shed light on cautionary tales. While the 15-Minute City concept is new, neighbourhood planning to improve society outcomes is not: in fact the literature has drawn parallels from the 15-Minute City to Clarence A. Perry's Neighbourhood Unit Concept (NUC) from the 1930s [@kissfazekasCircleParadigms15minute2022]. The NUC is a socio-spatial normative scheme widely adopted by government agencies (and the real-estate community) in the Western world after the Second World War [@talenSocialSciencePlanned2017; @solow1969concept]. Pairing well with the objectives of planning agencies at the time, Perry's NUC would allow for efficient mass-building of cellular units that prioritized the perceived functional needs of women and children: each unit providing proximate access to an elementary school and supporting community facilities, shopping, parks and housing [@talenSocialSciencePlanned2017; @brodyNeighbourhoodUnitConcept2013]. The NUC primarily prioritized local service provision, though Perry had confidence in good design's contribution to 'neighbourhood spirit' [@hall2014cities]. From the NUC's conception until the end of the 1960s, planners' aspirational attempts to prescribe social meaning to the neighbourhood's physical form had been criticized to near extinction [@talenSocialSciencePlanned2017]. By social scientists, the overestimation on the impact of built environment on social life was critiqued. By other planners, the inability to agree on the specificity of the neighbourhood (i.e., population size and the type, quality, and quantity of amenity) as well as how neighbourhood units are connected to other neighbourhoods and the rest of the region.  In response to these criticisms, neighbourhood planning proponents redefined deterministic terminology; their prescriptive physical form will 'enable' social contact with opportunities rather than 'engendering' it.

The re-defined bottom-up approach to community and neighbourhood planning has been adopted by New Urbanists, from which the 15-Minute City stems [@kissfazekasCircleParadigms15minute2022]; however, the effectiveness of these bottom-up processes remain to be seen due to the contemporary nature of the concept. Though, related research can be examined. In recent years, the question of _how can physical form be planned to enable an improved quality of life, and for whom with what outcomes_ has occupied the urban planning research agenda. For instance, the examination of low-income households access to transportation and their likelihood of gaining employment [@blumenbergDriveWorkRelationship2017; @bastiaanssenDoesBetterJob2022] or the relationship between children's access to public transit and participation in after-school activities [@palmRolePublicTransit2020]. A new wave of researchers and practitioners focused on local and context-specific relationships with the proximity to destinations have also emerged [@silvaProximitycentredAccessibilityConceptual2023; @silvaRegionalAccessibilityUndermining2022]. As reviewed  in the city plans that have adopted 15-Minute City approaches, the NUC's criticisms, and recent urban planning research, the question of _how to enable improved quality of life through urban built environment_ has shown to be highly context-sensitive, with the need for further investigation.

@@

In examining how to enable improved quality of life through urban built environment, an increasingly popular tool have been accessibility measures. These measures are a way to quantify the ease of reaching destinations from a given point in space and have been used to examine urban areas through just and sustainable city planning agendas [@vale2023accessibility; @handyAccessibilityIdeaWhose2020]. As the 15-Minute City is an amenity-provision neighbourhood planning concept which aims to enable urban environment that improves life quality, its analysis lends well to the use of accessibility tools [@guzmanProximityEnoughCritical2024]. Foreseeably, recent works have applied accessibility measures to investigate the 15-Minute City across various geographic scopes. For instance, in Naples, Italy [@gaglioneUrbanAccessibility15minute2022], Barcelona, Spain [@graellsCityCitiesMeasuring2021], Vancouver, Canada [@hosford15minuteCityReach2022], and in urban areas across Europe [@vale2023accessibility]. The "cumulative opportunity" measure has been applied to many 15-Minute City accessibility examinations. This form of accessibility quantifies how many destinations can be reached from a point in space within a given travel time threshold, pairing well with examinations that have a normative travel time threshold, like the 15-Minute City. Furthermore, the cumulative opportunity measure is widely appreciated for its intuitive computation and popularity among practitioners [@handyAccessibilityIdeaWhose2020; @handyMeasuringAccessibilityExploration1997; @chengInvestigatingWalkingAccessibility2019]. However, accessibility measures other than the cumulative opportunity have also been applied, reflecting the diversity of measures in the literature [@guzmanProximityEnoughCritical2024].

Another method which compliments the assessment of the 15-Minute City as it relates to the diversity of opportunity types, is entropy. The entropy measure has been widely used to characterise the diversity of land-use mix in early work by @frankLinkingObjectivelyMeasured2005 based on the species diversity Shannon-Wiener index that expresses relative evenness throughout a sequence [see review in @whittaker1972evolution]. It has also been used in social sciences to comprehend the occurrence of social events in a structured manner specially with regard to life trajectories [@ritschardSequenceAnalysisRelated2018]. In the realm of urban studies and planning literature, the malleable concept of entropy has been used to characterise the diversity in land-use mix [@ewingTravelBuiltEnvironment2010a], especially in the context of active modes [@luUrbanDensityDiversity2017; @mavoaIdentifyingAppropriateLanduse2018] and suburban sprawl [@randallGISBasedLand2015]. As well, it has been used to understand mobility behaviour [@mcbride2020exploration; @montero2023applying; @montero2023role]. Work that has used diversity indices alongside accessibility analysis are scarce, though examples of its use as parameters within accessibility scores in the case of employment opportunities [@chengMeasuringUrbanJob2013; @daiIncorporatingJobDiversity2018] and more recently to describe the diversity in transit facilities [@yinIncorporatingFacilityDiversity2024] are present. To the authors knowledge, the use of an entropy measure to reflect the diversity in care destination amenity types and in context within a 15-minute neighbourhood, has yet to be done. However, entropy measures theoretically compliment the 15-Minute City concept.

`r word_comment("I am only vaguely familar with this literature as of these months - how does this sound?", " ")`
In addition, classification algorithms have been useful tools in the identification of mobility and spatial typologies in relation to transportation planning. Machine learning approaches have often been deployed. For instance, @burkeGeospatialAnalysisFramework2022 uses spatially constrained multivariate clustering, a regionalisation technique, to develop urban form typologies related to the 15-Minute City. The k-means algorithm is another popular method; @ganUnderstandingUrbanMobility2020 applies it to analyse travel behaviour and classify metro stations based on mobility patterns. Combing classification techniques is also common; for instance, @delmelleDifferentiatingPathwaysNeighborhood2017 uses self-organizing maps (SOM) algorithm to group US neighbourhoods by minimising dissimilarity in attributes before applying k-means for further classification. Similarly, @victorianoTimeSpaceMoney2020 uses SOM to classify individuals' travel patterns according to the dissimilarity in mobility attributes and uses the decision tree algorithm to partition the data into interpretable classifications.

All-in-all, within the context of examining the 15-Minute City, a variety of tools have been applied in the literature. However, the question of _15-minute cities for whom_ sill demands further examination. This work aims to bridge the concept of the 15-Minute City with another normative framework: Mobility of Care.

@@

Rather than focusing on _all_ destinations, it may valuable to examine those related to caring activities. This re-framing offers a feminist perspective on  _what_ urban functions matter, connecting well with the 15-Minute City concept. Caring activities, which meet the physical, psychological, and emotional needs of others, are among the most essential and fundamental activities in society [@ilo_care_2018]. Yet, they are the most unequal, undervalued, and even devalued activities worldwide. Conventionally, caring activities have been borne on women’s shoulders [@haydenGrandDomesticRevolution1982; @hochschild_second_2012]. According to @ilo_care_2018, women and girls perform more than three-quarters of the total amount of unpaid care, a gender gap that varies geographically [@RN5]. This unequal share of caring responsibilities leads to multifaceted gendered differences: in career development, profession selection, contract type, pay gap, and time poverty, as recognized by various international organisations [@EIGE2016; @ilo_care_2018]. In terms of spatial and transportation planning, almost one third of daily trips are for care purposes [@sanchezdemadariagaMobilityCareIntroducing2013; @sanchezdemadariagaMeasuringMobilitiesCare2019; @ravensbergen2023exploratory]. From this research motivation, Sánchez de Madariaga coined the term Mobility of Care in 2013 to refer to all travel required to sustain the needs of a household, such as grocery shopping, escorting children, travelling to health appointments, and running errands [@sanchezdemadariagaMobilityCareIntroducing2013]. While an undercurrent of research had examined these unique household-serving trips over the decades, her seminal work was the first to consider all these trips as one category and demonstrate how mobility of care is a significant proportion of daily travel.

The Mobility of Care concept also explicitly integrates inter-sectional equity considerations. In considering Mobility of Care within the 15-Minute City one could address a common criticism leveled at the 15-Minute City [@guzmanProximityEnoughCritical2024]. Perhaps unsurprisingly given the gendered division of care work discussed, women have consistently been found to complete more mobility of care trips than men. In one study, mobility of care comprised 32% of women’s daily trips compared to 28% of men’s. While this gap is significant, it was found to be far greater in lower income households where women complete 20% more care trips than men [@ravensbergen2023exploratory]. Sánchez de Madariaga not only shows how important these mobility of care trips are, but also highlights the ways in which “mobility of care is systematically under-represented in any analysis of urban transport” (p. 37). As a product of the masculinst bias in planning, transport surveys and tools do not directly capture mobility of care, which re-enforces the idea that these trips are not a significant part of daily mobility. In this respect, the feminist perspective of the cities of proximity is still underestimated with only few examples on the topic [@gil_sola_choose_2022; @macintyre_her_2022].

Pairing well with the focus on shorter trips and the potential use of sustainable modes within the 15-Minute City, Mobility of Care trips are more local, shorter-distance and trip-chained. Compared to the trip to work, mobility of care trips are more frequently completed by foot, and less frequently by transit or bicycle [@ravensbergen2023exploratory]. However, little work to date examines walking to care destinations through the Mobility of Care framework. Though there is ample literature that examines walking to care destinations, such as schools (e.g., [@RN7;@RN8;@RN9]) and grocery stores (e.g., [@RN10;@RN11]), they tend to consider singular care destinations in research focusing on walkability.

The objectives of this research are two-fold. First, we aim to theoretically bridge Mobility of Care with the 15-Minute City concept to to re-imagine what local amenities matter and define an associated accessibility and diversity of accessibility measure. Second, taking a bottom-up data-driven approach, we apply a classification and clustering machine learning methodologies to support the bottom-up identification of _15-minute caring neighbourhoods_  through an empirical examination of the mid-cited city of Hamilton, Canada. These neighbourhoods are then interpreted on the continuum of having (and not having) the region-relative land-use and transportation infrastructure to support _completely caring_ access. Who currently resides in which neighbourhood is examined, and ways to further expand the care accessibility story are discussed.

`r word_comment("I need to clean this up and firmly reiterate the goals", " ")`
Due to it's popularity, the 15-Minute Concept is moving city planning towards a more people-centered approach focusing on accessibility to basic needs. While an aspirational ideal, the concept has faced criticism as a inadequately providing a way forward to address structural inequalities.  In this context, we bridge the 15-Minute City with the Mobility of Care concept, to add significance to what destinations and associated trips matter for all, but especially along gendered and socio-economic lines. We apply the "Caring 15-Minute City" concept to measure what neighourhoods in the mid-size Canadian city of Hamilton have the potential to be "caring 15 minute neighbourhoods". Making use of the tools reviewed and applied in the transportation research, the objectives of the paper are completed as follows:
